{\bf {\Large Title....}}

### 2016-07-22 ident_manu.Rnw
Wagener stuff, started structural uncertainty placeholder

A set of heuristics was developed to balance the tradeoff in model complexity and identifiability given the challenges described above.  These rulesets were developed with the assumption that parameters will be selected with preference for those with high sensitivity and identifability based on $\gamma < 15$ as an acceptable threshold for subsets (e.g., 93\% accountability).  Selection heurestics also recognized that parameter categories (i.e., optics, organic matter, phytoplankton, temperature, zooplankton) may have unequal preferences given questions of interest.  In all selection scenarios, parameters were selected by decreasing sensitivity starting with the most sensitive until identifiability did not exceed $\gamma = 15$ where selections were \begin{inparaenum}[1\upshape)]
\item blocked within parameter category,
\item independent of parameter category,
\item or considering all categories equally
\end{inparaenum}.  The selection rules produced seven subsets of parameters that could further be used to optimize model calibration for \ac{do}.

Finally, the above analyses were repeated for additional state variables estimated by FishTank to provide further descriptions of ecological dynamics that are relevant for hypoxia.  In addition to \ac{do}, other state variables included \ac{chla}, \ac{par}, nitrate, ammonium, particulate organic matter, dissolved organic matter, and phosphorus.  Particulate and dissolved organic matter were estimated as the summation of the respective outputs for organic matter from phytoplankon (\textit{OM1\_A}, \textit{OM2\_A}), fecal pellets (\textit{OM1\_fp}, \textit{OM2\_fp}), river sources \textit{OM1\_rp}, \textit{OM2\_rp}), and boundary conditions (\textit{OM1\_bc}, \textit{OM2\_bc}, see \citetalias{LehrterIR}).

@@

Response of \ac{do} to parameter perturbations was not uniform acros the time series.  \Cref{fig:sensplo} shows variation for the top parameters within each category.  Because FishTank does not include a spatial component, the estimated \ac{do} trend describes a closed, heterotrophic system where respiration processes eventually remove all \ac{do} from the model space.  The initial decrease in the time series reflects change from the initial conditions outside of the growing seasons (i.e., January, February), the spring/summer increase represents production associated with expected seasonal growth, and the remaining time series from August to the end of year shows complete removal of \ac{do} as respiration processes dominate metabolic activity. Although this time series is not a realistic depiction of an actual system, the biogeochemical model behaves as expected in the absence of the hydrodynamic model.  Accordingly, the interpretation of sensitivity results from the simple model has relevance in an ecological context.  As expected, parameter perturbations had the largest effect during the summer months, although the effects varied.  An inrease in 50\% from the parameter default values generally caused a reduction in \ac{do} during the summer, with the exception of the zooplankton parameter, \Sexpr{sensrngcat$Zooplankton$Parameter[2]}, which caused an increase in \ac{do}.  The phytoplankton parameter, \Sexpr{sensrngcat$Phytoplankton$Parameter[2]}, had the largest effect such that the \ac{do} time series was similar to the default output in April/May, whereas a dramatic decrease was observed in the remaining months.  The effects of perturbations early in the time series (January, February) showed similar patterns such that a reduction in \ac{do} was most common, particularly for the optics (\Sexpr{sensrngcat$Optics$Parameter[2]}) and phytoplankton (\Sexpr{sensrngcat$Phytoplankton$Parameter[2]}) parameters.

@@

@@

How does the assimilation of additional parameters (e.g., other state variables) during calibration influence the conclusions?

@@

>>>>

A set of heuristics was developed to balance the tradeoff in model complexity and identifiability given the challenges described above \citep[see also][]{Wagener01}.  These rulesets were developed with the assumption that parameters will be selected with preference for those with high sensitivity and identifability based on $\gamma < 15$ as an acceptable threshold for subsets (e.g., 93\% accountability).  Selection heurestics also recognized that parameter categories (i.e., optics, organic matter, phytoplankton, temperature, zooplankton) may have unequal preferences given questions of interest.  In all selection scenarios, parameters were selected by decreasing sensitivity starting with the most sensitive until identifiability did not exceed $\gamma = 15$ where selections were \begin{inparaenum}[1\upshape)]
\item blocked within parameter category,
\item independent of parameter category,
\item or considering all categories equally
\end{inparaenum}.  The selection rules produced seven subsets of parameters that could further be used to optimize model calibration for \ac{do}.

The above analyses were repeated for additional state variables estimated by FishTank to provide further descriptions of ecological dynamics that are relevant for hypoxia.  In addition to \ac{do}, other state variables included \ac{chla}, \ac{par}, nitrate, ammonium, particulate organic matter, dissolved organic matter, and phosphorus.  Particulate and dissolved organic matter were estimated as the summation of the respective outputs for organic matter from phytoplankon (\textit{OM1\_A}, \textit{OM2\_A}), fecal pellets (\textit{OM1\_fp}, \textit{OM2\_fp}), river sources \textit{OM1\_rp}, \textit{OM2\_rp}), and boundary conditions (\textit{OM1\_bc}, \textit{OM2\_bc}, see \citetalias{LehrterIR}).

\subsection{Identifiability and structural uncertainty}

The sensitivity and identifiability analyses can be extended to

@@

Response of \ac{do} to parameter perturbations was not uniform across the time series.  \Cref{fig:sensplo} shows variation for the top parameters within each category.  Because FishTank does not include a spatial component, the estimated \ac{do} trend describes a closed, heterotrophic system where respiration processes eventually remove all \ac{do} from the model space.  The initial decrease in the time series reflects change from the initial conditions outside of the growing seasons (i.e., January, February), the spring/summer increase represents production associated with expected seasonal growth, and the remaining time series from August to the end of year shows complete removal of \ac{do} as respiration processes dominate metabolic activity. Although this time series is not a realistic depiction of an actual system, the biogeochemical model behaves as expected in the absence of the hydrodynamic model.  Accordingly, the interpretation of sensitivity results from the simple model has relevance in an ecological context.  As expected, parameter perturbations had the largest effect during the summer months, although the effects varied.  An inrease in 50\% from the parameter default values generally caused a reduction in \ac{do} during the summer, with the exception of the zooplankton parameter, \Sexpr{sensrngcat$Zooplankton$Parameter[2]}, which caused an increase in \ac{do}.  The phytoplankton parameter, \Sexpr{sensrngcat$Phytoplankton$Parameter[2]}, had the largest effect such that the \ac{do} time series was similar to the default output in April/May, whereas a dramatic decrease was observed in the remaining months.  The effects of perturbations early in the time series (January, February) showed similar patterns such that a reduction in \ac{do} was most common, particularly for the optics (\Sexpr{sensrngcat$Optics$Parameter[2]}) and phytoplankton (\Sexpr{sensrngcat$Phytoplankton$Parameter[2]}) parameters.

@@

We showed only small subsets are identifiable, similar conclusions have been described by citations in \citep{Wagener01}, p. 14 for models that follow traditional calibration schemes (e.g., objective function minimization).

@@

How does the assimilation of additional parameters (e.g., other state variables) during calibration influence the conclusions?  \citet{Wagener01} describes this as a potential approach to improving model performance by improving the availability of information for model calibration (p. 14).

@@

Our heuristic products are partially analogous to the Rainfall-Runoff Modelling Toolbox (RRMT) presented by Wagener et al. 1999, 2001 (cited on page 15, in \citealt{Wagener01}).

### 2016-07-22 ident_manu.tex
Wagener stuff, started structural uncertainty placeholder

\begin{singlespace}
\title{{\bf {\Large Title....}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, John C. Lehrter$^1$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:lehrter.john@epa.gov}{lehrter.john@epa.gov}}}
  \vspace{1in} 
  \\ Version Date:   Mon Jul 11 10:46:03 2016 -0500
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

A set of heuristics was developed to balance the tradeoff in model complexity and identifiability given the challenges described above.  These rulesets were developed with the assumption that parameters will be selected with preference for those with high sensitivity and identifability based on $\gamma < 15$ as an acceptable threshold for subsets (e.g., 93\% accountability).  Selection heurestics also recognized that parameter categories (i.e., optics, organic matter, phytoplankton, temperature, zooplankton) may have unequal preferences given questions of interest.  In all selection scenarios, parameters were selected by decreasing sensitivity starting with the most sensitive until identifiability did not exceed $\gamma = 15$ where selections were \begin{inparaenum}[1\upshape)]
\item blocked within parameter category,
\item independent of parameter category,
\item or considering all categories equally
\end{inparaenum}.  The selection rules produced seven subsets of parameters that could further be used to optimize model calibration for \ac{do}.

Finally, the above analyses were repeated for additional state variables estimated by FishTank to provide further descriptions of ecological dynamics that are relevant for hypoxia.  In addition to \ac{do}, other state variables included \ac{chla}, \ac{par}, nitrate, ammonium, particulate organic matter, dissolved organic matter, and phosphorus.  Particulate and dissolved organic matter were estimated as the summation of the respective outputs for organic matter from phytoplankon (\textit{OM1\_A}, \textit{OM2\_A}), fecal pellets (\textit{OM1\_fp}, \textit{OM2\_fp}), river sources \textit{OM1\_rp}, \textit{OM2\_rp}), and boundary conditions (\textit{OM1\_bc}, \textit{OM2\_bc}, see \citetalias{LehrterIR}).

@@

Response of \ac{do} to parameter perturbations was not uniform acros the time series.  \Cref{fig:sensplo} shows variation for the top parameters within each category.  Because FishTank does not include a spatial component, the estimated \ac{do} trend describes a closed, heterotrophic system where respiration processes eventually remove all \ac{do} from the model space.  The initial decrease in the time series reflects change from the initial conditions outside of the growing seasons (i.e., January, February), the spring/summer increase represents production associated with expected seasonal growth, and the remaining time series from August to the end of year shows complete removal of \ac{do} as respiration processes dominate metabolic activity. Although this time series is not a realistic depiction of an actual system, the biogeochemical model behaves as expected in the absence of the hydrodynamic model.  Accordingly, the interpretation of sensitivity results from the simple model has relevance in an ecological context.  As expected, parameter perturbations had the largest effect during the summer months, although the effects varied.  An inrease in 50\% from the parameter default values generally caused a reduction in \ac{do} during the summer, with the exception of the zooplankton parameter, \textit{ZKa$_{z1}$}, which caused an increase in \ac{do}.  The phytoplankton parameter, \textit{Qc$_{p1}$}, had the largest effect such that the \ac{do} time series was similar to the default output in April/May, whereas a dramatic decrease was observed in the remaining months.  The effects of perturbations early in the time series (January, February) showed similar patterns such that a reduction in \ac{do} was most common, particularly for the optics (\textit{astar490}) and phytoplankton (\textit{Qc$_{p1}$}) parameters.

>>>>

\begin{singlespace}
\title{{\bf {\Large Title....}}}
\author{
  {\bf {\normalsize Marcus W. Beck$^1$, John C. Lehrter$^1$}}
  \\\\{\textit {\normalsize $^1$USEPA National Health and Environmental Effects Research Laboratory}}
  \\{\textit {\normalsize Gulf Ecology Division, 1 Sabine Island Drive, Gulf Breeze, FL 32561}}
	\\{\textit {\normalsize Phone: 850-934-2480, Fax: 850-934-2401}}
	\\{\textit {\normalsize Emails: \href{mailto:beck.marcus@epa.gov}{beck.marcus@epa.gov}, \href{mailto:lehrter.john@epa.gov}{lehrter.john@epa.gov}}}
  \vspace{1in} 
  \\ Version Date:   Mon Jul 11 11:13:58 2016 -0500
	}
\date{}
\maketitle
\end{singlespace}
\clearpage

@@

A set of heuristics was developed to balance the tradeoff in model complexity and identifiability given the challenges described above \citep[see also][]{Wagener01}.  These rulesets were developed with the assumption that parameters will be selected with preference for those with high sensitivity and identifability based on $\gamma < 15$ as an acceptable threshold for subsets (e.g., 93\% accountability).  Selection heurestics also recognized that parameter categories (i.e., optics, organic matter, phytoplankton, temperature, zooplankton) may have unequal preferences given questions of interest.  In all selection scenarios, parameters were selected by decreasing sensitivity starting with the most sensitive until identifiability did not exceed $\gamma = 15$ where selections were \begin{inparaenum}[1\upshape)]
\item blocked within parameter category,
\item independent of parameter category,
\item or considering all categories equally
\end{inparaenum}.  The selection rules produced seven subsets of parameters that could further be used to optimize model calibration for \ac{do}.

The above analyses were repeated for additional state variables estimated by FishTank to provide further descriptions of ecological dynamics that are relevant for hypoxia.  In addition to \ac{do}, other state variables included \ac{chla}, \ac{par}, nitrate, ammonium, particulate organic matter, dissolved organic matter, and phosphorus.  Particulate and dissolved organic matter were estimated as the summation of the respective outputs for organic matter from phytoplankon (\textit{OM1\_A}, \textit{OM2\_A}), fecal pellets (\textit{OM1\_fp}, \textit{OM2\_fp}), river sources \textit{OM1\_rp}, \textit{OM2\_rp}), and boundary conditions (\textit{OM1\_bc}, \textit{OM2\_bc}, see \citetalias{LehrterIR}).

\subsection{Identifiability to evaluate structural uncertainty}

The sensitivity and identifiability analyses can be extended to

@@

Response of \ac{do} to parameter perturbations was not uniform across the time series.  \Cref{fig:sensplo} shows variation for the top parameters within each category.  Because FishTank does not include a spatial component, the estimated \ac{do} trend describes a closed, heterotrophic system where respiration processes eventually remove all \ac{do} from the model space.  The initial decrease in the time series reflects change from the initial conditions outside of the growing seasons (i.e., January, February), the spring/summer increase represents production associated with expected seasonal growth, and the remaining time series from August to the end of year shows complete removal of \ac{do} as respiration processes dominate metabolic activity. Although this time series is not a realistic depiction of an actual system, the biogeochemical model behaves as expected in the absence of the hydrodynamic model.  Accordingly, the interpretation of sensitivity results from the simple model has relevance in an ecological context.  As expected, parameter perturbations had the largest effect during the summer months, although the effects varied.  An inrease in 50\% from the parameter default values generally caused a reduction in \ac{do} during the summer, with the exception of the zooplankton parameter, \textit{ZKa$_{z1}$}, which caused an increase in \ac{do}.  The phytoplankton parameter, \textit{Qc$_{p1}$}, had the largest effect such that the \ac{do} time series was similar to the default output in April/May, whereas a dramatic decrease was observed in the remaining months.  The effects of perturbations early in the time series (January, February) showed similar patterns such that a reduction in \ac{do} was most common, particularly for the optics (\textit{astar490}) and phytoplankton (\textit{Qc$_{p1}$}) parameters.
